Local Spectral Clustering of Density Upper Level Sets

### 2019-11-07 arxiv/local_spectral_density_clustering_arxiv.tex
siva's updates

{\large{
\begin{center}
Alden Green ~~~~~ Sivaraman Balakrishnan~~~~~ Ryan Tibshirani\\
\vspace{.2cm}
\end{center}

@@

\begin{abstract}
\noindent We analyze the Personalized PageRank (PPR) algorithm, a local spectral method for clustering,
which extracts clusters using locally-biased random walks around a user-specified
seed node.  In contrast to previous work, we adopt a traditional statistical
learning setup, where we obtain samples from an unknown distribution, and aim to
identify connected regions of high-density (density clusters).  We prove that
PPR, run on a neighborhood graph, extracts sufficiently salient density
clusters, and conversely fails to recover geometrically poorly conditioned density clusters. We also provide empirical support to illustrate our theoretical results.
\end{abstract}
\end{center}
\section{Introduction}
\label{sec: introduction}
In this paper, we consider the problem of clustering: splitting a given data set
into groups that satisfy some notion of within-group similarity and
between-group difference.  Our particular focus is on
spectral clustering methods, a family of powerful
nonparametric clustering algorithms. Roughly speaking, a spectral algorithm
first constructs a geometric graph $G$, where vertices correspond to samples,
and edges correspond to proximities between samples. The algorithm 
then estimates a feature
embedding based on (an appropriate) Laplacian matrix of the graph 
$G$, and applies a simple clustering
technique (like $k$-means clustering) in the embedded feature space.

@@

When applied to geometric graphs built from a large number of samples,
global spectral clustering methods can be computationally cumbersome and   
insensitive to the local geometry of the underlying distribution
\citep{leskovec2010,mahoney2012}.  This has led to increased interest in
local spectral clustering algorithms, which leverage locally-biased spectra
computed using random walks around some user-specified seed node.  A popular 
local clustering algorithm is the Personalized PageRank (PPR) algorithm, first introduced by 
\citet{haveliwala2003}, then further developed by
several others~\citep{spielman2011,spielman2014,andersen2006,mahoney2012,zhu2013}.

Local spectral clustering techniques have been practically very successful
\citep{leskovec2010,andersen2012,gleich2012,mahoney2012,wu2012}, leading 
many authors to develop supporting theory
\citep{spielman2013,andersen2009,gharan2012,zhu2013} that gives worst-case
guarantees on traditional graph-theoretic notions of cluster quality (such as
conductance).  In this paper, we adopt a classical statistical viewpoint,
and examine what the output of local clustering on a data set reveals about the
underlying density of the samples 
$f$.  In particular, we examine the ability of PPR to recover
\emph{density clusters} of $f$, defined as the connected components of
the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$ for some $\lambda > 0$
(a central object of interest in the statistical clustering literature, dating
back to the work of \citet{hartigan1981}).

@@

\subsection{Sweep cut.}
\label{subsection: sweep_cut}

We employ the sweep cut method to obtain a local cluster from the PPR embedding $\pbf_v$. For a level $\beta > 0$, we define a \emph{$\beta$-sweep cut} of $\pbf_v = (\pbf_v(u))_{u \in V}$
as   
\begin{equation}
\label{eqn: sweep_cuts}
S_{\beta,v} := \set{u \in V: \frac{\pbf_v(u)}{\Dbf_{uu}} > \beta}.
\end{equation}
We will use the normalized cut metric to determine which sweep cut $S_{\beta}$
is the best cluster estimate. For a set $S \subseteq V$ with complement $S^c = V
\setminus S$, we define \smash{$\cut(S;G_{n,r}) := \sum_{u \in S, v \in S^c}
\Abf_{uv}$}, and \smash{$\vol(S; G_{n,r}) := \sum_{u \in S} \Dbf_{uu}$}.  We
define the \emph{normalized cut} of $S$ as
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
Having computed sweep cuts $S_{\beta}$ over a range \smash{$\beta \in  
(L, U)$} (where the range $(L,U)$ is also a user-specified parameter), we output the cluster estimate
\smash{$\Cest = S_{\beta^*}$} that has minimum normalized cut.
%\smash{$\Phi(S_{\beta^*}; G_{n,r})$}. 
For concreteness, the PPR algorithm is summarized in Algorithm~\ref{alg: ppr}.

@@

\subsection{Related Work and Our Contributions}
In addition to the background on local spectral
clustering given previously, a few related lines of work are worth
highlighting. In the stochastic block model (SBM), arguably one of the simplest models of network formation, edges between nodes independently occur with probability based on a latent community membership. The ability of spectral algorithms to perform clustering -- community detection-- in this context is well-understood. In particular, \citet{rohe2011} upper bound the fraction of nodes misclassified by a spectral algorithm for the high-dimensional (large number of blocks) SBM,  \citet{lei2015} extend these results to the sparse (low average degree) regime, and \citet{balakrishnan2011} analyze the misclassification rate when the block model exhibits some hierarchical structure. The framework we consider, where nodes correspond to data points sampled from an underlying density, and edges between nodes are formed based on geometric proximity, is quite different than the SBM, and therefore so is our analysis.

@@

A substantial portion of our technical effort involves estimating graph-theoretic connectivity properties of empirical density clusters in the neighborhood graph $G_{n,r}$. In particular, we prove an upper bound on the mixing time of a random walk run only over the subset of nodes in $G_{n,r}$ which fall within a density cluster $\Cset$. To do so, we rely on a series of seminal works upper bounding the mixing time of \emph{geometric random walks} (see \citet{vempala2005} for a comprehensive review.)  This study was initiated by \citet{dyer1991}, who used geometric random walks as a fundamental subroutine to efficiently compute volumes of high-dimensional convex bodies. These results are improved in \citet{lovasz1990,kannan97,kannan06}, who show \textit{inter alia} that the bounds on mixing time can be sharpened by avoiding so-called ``start-penalties''.(These improvements are crucial to our work, for they allow us to avoid factors of $\log n$ which would imply divergence as $n \to \infty$.) Following the lead of \citet{abbasi-yadkori2016,abbasi-yadkori2016a}, we extend these results to hold for Lipschitz deformations of convex sets. Additionally, we relate the mixing time of these (continuous-space) geometric random walks to the mixing time of random walks over (discrete) neighborhood graphs.

Finally, it is worth mentioning that density clustering and level set estimation are well-studied problems. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013} consider consistent estimation of the
cluster tree.
%, to note but a few works. 
We emphasize that our goal is not to improve on these
results, nor to offer a better algorithm for level set estimation; indeed, seen as
a density clustering algorithm, PPR has none of the optimality guarantees 
found in the aforementioned works. Instead, our motivation is to start with a 
widely-used local spectral method, PPR, and to better understand and
characterize the distinctions between those density clusters which are
well-conditioned for PPR, and those which are not.

@@

\item We further show in Section \ref{sec:
    consistent_cluster_estimation_with_ppr} that if the density cluster 
  $\Cset$ is particularly well-conditioned, Algorithm \ref{alg: ppr}
  will consistently estimate a density cluster in the sense of
  \eqref{eqn: consistent_density_cluster_recovery}.

\item In Section \ref{sec: analysis}, we detail some of the analysis required to
  prove our main results, and expose the parts various geometric quantities play 
  in the difficulty of the clustering problem.

@@

\noindent At this point we summarize one of our main takeaways: 
PPR, run on a neighborhood graph, recovers only \emph{geometrically
compact} high-density clusters. Our main theoretical results, make this takeaway 
precise, and provide concrete ways of quantifying the geometric
compactness of a density cluster.

@@

  We assume that neighborhood graph radius $0 < r \leq \sigma/2d$ is chosen such that
  \begin{equation*}
  \vol_{\Pbb,r}(\mathcal{S}) \leq \frac{1}{2}\vol_{\Pbb,r}(\Reals^d).
  \end{equation*}

@@

\subsection{Cluster Recovery in Symmetric Set Difference}
We now present our first main result: a bound on the volume of the symmetric set difference between the estimated cluster $\widehat{C}$ and empirical cluster $\Csig[\Xbf]$. 
\begin{theorem}
  \label{thm:volume_ssd}
  Fix $\lambda > 0$ and let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster for some $\sigma > 0$. If Algorithm \ref{alg: ppr} is \emph{well-initialized} with respect to $\Cset$, then there exists a universal constant $c > 0$, constants in sample size $c_1(\Pbb,r)$ - $c_6(\Pbb,r)$, and a set $\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$ with $\vol(\Csig[\Xbf]^g; G_{n,r}) \geq \frac{1}{2}\vol(\Csig[\Xbf];G_{n,r})$ such that the following statement holds: let Algorithm~\ref{alg: ppr} be run with any seed node $v \in \Csig[\Xbf]^g$. Then, for any $n$ large enough so that
  \begin{equation*}
  n \geq \max\left\{(\log n)^{\max\{\frac{3}{d},1\}}\left(\frac{1}{c_2(\Pbb,r)}\right)^d, c_3(\Pbb,r)\right\},
  \end{equation*}
  the volume of the symmetric set difference is upper bounded
  \begin{equation}
    \label{eqn:volume_ssd}
    \Delta(\Csig[\Xbf], \Cest) \leq c \cdot \kappa(\Cset) \cdot \vol_{n,r}(\Csig[\Xbf]),
  \end{equation}
  with probability at least $1 - \frac{c_4(\Pbb,r)}{n} - 2n\exp\set{-c_5(\Pbb,r)n} - 2\exp\set{-(c_1(\Pbb,r) + c_6(\Pbb,r))n}$. 
\end{theorem}
The proof of Theorem \ref{thm:volume_ssd} is deferred to Appendix~\ref{sec:proof_of_volume_ssd}. In this theorem, and hereafter, we let $c, c_i > 0$ represent universal constants, and $c_i(\Pbb,r)$ represent constants which depend only on $\Pbb$ and $r$ but not on the sample size $n$; where possible, we keep track of all constants in our proofs. 
This result shows that the volume of the symmetric set difference \smash{$\Delta(\Csig[\Xbf],
  \Cest)$} is upper-bounded by a quantity proportional to the difficulty of the clustering problem, as
measured by the condition number $\kappa(\Cset)$.

\subsection{Consistent Cluster Recovery}
The bound on symmetric set difference \eqref{eqn:volume_ssd} does not imply consistent density cluster estimation in the sense of \eqref{eqn: consistent_density_cluster_recovery}. This notion of consistency requires a uniform bound over $\pbf$: as an example, for all \smash{$\Cset'
  \in \Cbb_f(\lambda), \Cset' \neq \Cset$}, and each $u \in \Cset, w \in
\Cset'$,  
\begin{equation}
\label{eqn: ppr_gap}
\frac{\pbf_v(w)}{D_{ww}} \leq \frac{1}{100 {n \choose 2} \vol_{\Pbb,r}(\Csig)} < \frac{1}{10 {n \choose 2} \vol_{\Pbb,r}(\Csig)} \leq \frac{\pbf_v(u)}{D_{uu}}.
\end{equation}
Then, any $(L,U)$ satisfying~\eqref{eqn: initialization} and any sweep cut $S_{\beta}$ for $\beta \in (L,U)$ will fulfill both conditions
laid out in \eqref{eqn: consistent_density_cluster_recovery}. In Theorem
\ref{thm: consistent_recovery_of_density_clusters}, we show that a sufficiently
small upper bound on $\kappa(\Cset)$ ensures that with high probability the uniform bound~\eqref{eqn: ppr_gap} is satisfied, and hence implies \smash{$\Cest$} will be a consistent  
estimator. We will need one additional regularity condition, to preclude arbitrarily low degree
vertices for points $x \in \Cset'[\Xbf]$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{6}
\item 
  \label{asmp: C'_bounded_density}
  \emph{Bounded density in other clusters:} Letting $\sigma,\lambda_{\sigma}$ be 
  as in \ref{asmp: bounded_density}, for each $\Cset' \in \Cbb_f(\lambda)$ and
  for all $x \in \Csig'$, $\lambda_{\sigma} \leq f(x)$. 
\end{enumerate}

@@

\begin{theorem}
  \label{thm: consistent_recovery_of_density_clusters}
  Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a
  $\kappa$-well-conditioned density cluster for some $\sigma > 0$, and
  additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >  0$ and constants in sample size $c_1(\Pbb,r)$-$c_6(\Pbb,r)$ such that if  
  \begin{equation}
    \label{eqn: kappa_ub}
    \kappa(\Cset) \leq c \frac{(\lambda_{\sigma} r^d
      \nu_d)^2}{\vol_{\Pbb,r}(\Csig)}, 
  \end{equation}
  then $\Cest$ satisfies~\eqref{eqn: consistent_density_cluster_recovery} with respect to $\Cset$ with probability at least $1 - \frac{c_4(\Pbb,r)}{n} - 2n\exp\set{-c_5(\Pbb,r)n} - 2\exp\set{-(c_1(\Pbb,r) + c_6(\Pbb,r))n}$. Therefore, $\Cest$ is a consistent estimator of $\Cset$ in the sense of Definition \ref{def: consistent_density_cluster_estimation}. 
\end{theorem}
\noindent Some remarks are in order: 
 \begin{enumerate}  
 \item We note that the restriction on $\kappa(\Cset)$ imposed by \eqref{eqn:
    kappa_ub} results in an upper bound on the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf],
    \Cest)$} on the order of $r^d$. In plain terms, we are able to recover a
  density cluster $\Cset$ in the sense of \eqref{eqn:
    consistent_density_cluster_recovery} only when we can guarantee a very small
  fraction of points will be misclassified. This strong condition is the price
  we pay in order to obtain the uniform bound in~\eqref{eqn: ppr_gap}. 
 \item While taking the radius of the neighborhood graph $r \to 0$ as $n \to
  \infty$ (thereby ensuring that the graph $G_{n,r}$ is sparse) is computationally
  attractive, the presence of a factor of \smash{$\log^2(1/r)/r$} in
  $\kappa(\Cset)$ prevents us from making claims about the
  behavior of PPR in this regime. Although the restriction to a kernel
  function fixed in $n$ is common in spectral clustering theory
  \citep{schiebinger2015,vonluxburg2008}, it is an interesting question whether
  PPR exhibits some degeneracy over $r$-neighborhood graphs as $r \to 0$, or if
  this is merely looseness in our upper bounds.  
\end{enumerate}

\subsection{Approximate PPR vector} 
In practice, exactly solving the system of equations~\eqref{eqn:
  ppr_vector} to compute the PPR vector 
  may be too computationally expensive. To address this limitation,
\citet{andersen2006} introduced the \emph{$\varepsilon$-approximate} PPR vector
(aPPR), which we will denote by \smash{$\pbf^{(\varepsilon)}$}. We refer the
curious reader to \citet{andersen2006} for a formal algorithmic definition of
the aPPR vector, and limit ourselves to highlighting a few salient points: the
aPPR vector can be computed in order $\mathcal{O}(1/(\varepsilon\alpha))$ time,
while satisfying the following uniform error bound: 
\begin{equation}
\label{eqn: appr_error}
\textrm{for all $u \in V$}, \quad \pbf(u) - \varepsilon \Dbf_{uu}\leq
\pbf^{(\varepsilon)}(u) \leq \pbf(u).  
\end{equation}
For a sufficiently small choice of $\varepsilon$, the 
application of \eqref{eqn: appr_error} within the proofs of Theorems \ref{thm:volume_ssd} and \ref{thm: consistent_recovery_of_density_clusters}
leads to analogous results which hold for \smash{$\pbf^{(\varepsilon)}$}. We
formally state and prove this in Corollary~\ref{cor: appr} in 
Appendix~\ref{sec:appr}.

@@

One of the primary technical contributions of our work is showing that the geometric
conditions \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} translate to 
meaningful bounds on the normalized cut and mixing time of $\Csig[\Xbf]$ in
$G_{n,r}$. In doing so, we elaborate on how some of the geometric conditions
introduced in Section \ref{sec: consistent_cluster_estimation_with_ppr}
contribute to the difficulty of the clustering problem.

@@

\begin{remark}Technically, the rectangles $\Csig^{(1)}$ and $\Csig^{(2)}$ are not $\sigma$-expansions due to their sharp corners. To fix this, one can either modify the upper bound (specifically, Lemma~\ref{lem: expansion_volume} in the proof of Theorem~\ref{thm: conductance_upper_bound}) to hold with respect to rectangles of width $\sigma$, or change the sets $\Csig^{(1)}$ and $\Csig^{(2)}$ to be rectangles with rounded corners. As these details do not meaningfully change our conclusions, for simplicity we will keep our upper and lower bounds as is.
\end{remark}

@@

>>>>

{\large{
\begin{center}
Alden Green~~~~~ Sivaraman Balakrishnan~~~~~ Ryan Tibshirani\\
\vspace{.2cm}
\end{center}

@@

\begin{abstract}
\noindent We analyze the Personalized PageRank (PPR) algorithm, a local spectral method for clustering,
which extracts clusters using locally-biased random walks around a user-specified
seed node.  In contrast to previous work, we adopt a classical statistical
learning setup, where we obtain samples from an unknown distribution, and aim to
identify connected regions of high-density (density clusters).  We prove that
PPR, run on a neighborhood graph, extracts sufficiently salient density
clusters. We also show a surprising converse result, that PPR can fail 
to recover geometrically poorly conditioned density clusters, even asymptotically. Finally, we provide empirical support to illustrate our theoretical results.
\end{abstract}
\end{center}
\section{Introduction}
\label{sec: introduction}
In this paper, we consider the problem of clustering: splitting a given data set
into groups that satisfy some notion of within-group similarity and
between-group difference.  Our particular focus is on
spectral clustering methods, a family of powerful
nonparametric clustering algorithms. Roughly speaking, a spectral algorithm
first constructs a geometric graph $G$, where vertices correspond to samples,
and edges correspond to proximities between samples. The algorithm 
then estimates a feature
embedding based on (an appropriate) Laplacian matrix of the graph 
$G$, and applies a simple clustering
technique (like $k$-means clustering) in the embedded feature~space.

@@

When applied to geometric graphs built from a large number of samples,
global spectral clustering methods can be computationally cumbersome and   
insensitive to the local geometry of the underlying distribution
\citep{leskovec2010,mahoney2012}.  This has led to increased interest in
\emph{local} spectral clustering algorithms, which leverage locally-biased spectra
computed using random walks around some user-specified seed node.  A popular 
local clustering algorithm is the Personalized PageRank (PPR) algorithm, first introduced by 
\citet{haveliwala2003}, then further developed by
several others~\citep{spielman2011,spielman2014,andersen2006,mahoney2012,zhu2013}.

Local spectral clustering techniques have been practically very successful
\citep{leskovec2010,andersen2012,gleich2012,mahoney2012,wu2012}, leading 
many authors to develop supporting theory
\citep{spielman2013,andersen2009,gharan2012,zhu2013} that gives worst-case
guarantees on traditional graph-theoretic notions of cluster quality (such as
conductance).  In this paper, we adopt a classical statistical viewpoint,
and examine what the output of local clustering on a data set reveals about the
underlying density $f$ of the samples.  In particular, we examine the ability of PPR to recover
\emph{density clusters} of $f$, defined as the connected components of
the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$ for some $\lambda > 0$
(a central object of interest in the statistical clustering literature, dating
back to the work of \citet{hartigan1981}).

@@

\subsubsection{Sweep cut}
\label{subsection: sweep_cut}
We employ the sweep cut method to obtain a local cluster from the PPR embedding $\pbf_v$. In more detail, for a level $\beta > 0$, we define a \emph{$\beta$-sweep cut} of $\pbf_v = (\pbf_v(u))_{u \in V}$
as   
\begin{equation}
\label{eqn: sweep_cuts}
S_{\beta,v} := \set{u \in V: \frac{\pbf_v(u)}{\Dbf_{uu}} > \beta}.
\end{equation}
We will use the normalized cut metric to determine which sweep cut $S_{\beta}$
is the best cluster estimate. For a set $S \subseteq V$ with complement $S^c = V
\setminus S$, we define \smash{$\cut(S;G_{n,r}) := \sum_{u \in S, v \in S^c}
\Abf_{uv}$}, and \smash{$\vol(S; G_{n,r}) := \sum_{u \in S} \Dbf_{uu}$}.  We
define the \emph{normalized cut} of $S$ as
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
Having computed sweep cuts $S_{\beta}$ over a range \smash{$\beta \in  
(L, U)$} (where the range $(L,U)$ is also a user-specified parameter), we output the cluster estimate
\smash{$\Cest = S_{\beta^*}$} that has minimum normalized cut.
%\smash{$\Phi(S_{\beta^*}; G_{n,r})$}. 
For concreteness, the PPR algorithm is summarized in Algorithm~\ref{alg: ppr}.

@@

\subsection{Related Work and Our Contributions}
In addition to the background on local spectral
clustering given previously, a few related lines of work are worth
highlighting. In the stochastic block model (SBM), arguably one of the simplest models of network formation, edges between nodes independently occur with probability based on a latent community membership. The ability of spectral algorithms to perform clustering -- community detection-- in this context is well-understood. \sbcomment{Add reference to McSherry -- spectral partitioning...} In particular, \citet{rohe2011} upper bound the fraction of nodes misclassified by a spectral algorithm for the high-dimensional (large number of blocks) SBM,  \citet{lei2015} extend these results to the sparse (low average degree) regime, and \citet{balakrishnan2011} analyze the misclassification rate when the block model exhibits some hierarchical structure
\sbcomment{add reference to Cris Moore, Newman et. al., and Liza Levina, Bickel, Sarkar have a recent paper that looks at this same hierarchical model}. The framework we consider, where nodes correspond to data points sampled from an underlying density, and edges between nodes are formed based on geometric proximity, is quite different than the SBM, and therefore so is our analysis.

@@

A substantial portion of our technical effort involves estimating graph-theoretic connectivity properties of empirical density clusters in the neighborhood graph $G_{n,r}$. In particular, we prove an upper bound on the mixing time of a random walk run only over the subset of nodes in $G_{n,r}$ which fall within a density cluster $\Cset$. To do so, we rely on a series of seminal works upper bounding the mixing time of \emph{geometric random walks} (see \citet{vempala2005} for a comprehensive review.)  This study was initiated by \citet{dyer1991}, who used geometric random walks as a fundamental subroutine to efficiently compute volumes of high-dimensional convex bodies. These results are improved in \citet{lovasz1990,kannan97,kannan06}, who show, inter alia, that the bounds on mixing time can be sharpened by avoiding so-called ``start-penalties''. As we discuss further in the sequel, these improvements are crucial to our work. Following the work of \citet{abbasi-yadkori2016,abbasi-yadkori2016a}, we extend these results to hold for Lipschitz deformations of convex sets. Additionally, we relate the mixing time of these (continuous-space) geometric random walks to the mixing time of random walks over (discrete) neighborhood graphs. \sbcomment{Refer to and add a couple of lines about the paper of Zhu et al.}

Finally, it is worth mentioning that density clustering and level set estimation are well-studied problems. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013} \sbcomment{Add reference to Kpotufe, von Luxburg} consider consistent estimation of the
cluster tree.
%, to note but a few works. 
We emphasize that our goal is not to improve on these
results, nor to offer a better algorithm for level set estimation; indeed, seen as
a density clustering algorithm, PPR has none of the optimality guarantees 
found in the aforementioned works. Instead, our motivation is to start with a 
widely-used local spectral method, PPR, and to better understand and
characterize the distinctions between those density clusters which are
well-conditioned for PPR, and those which are not.

@@

\item We further show in Section \ref{sec:
    consistent_cluster_estimation_with_ppr} that if the density cluster 
  $\Cset$ is particularly well-conditioned, Algorithm \ref{alg: ppr}
  will consistently estimate a density cluster in a strong sense (see~\eqref{eqn: consistent_density_cluster_recovery}).

\item In Section \ref{sec: analysis}, we detail some of the analysis required to
  prove our main results, and expose the roles that various geometric quantities play 
  in the difficulty of the clustering problem.

@@

\noindent At this point we summarize one of our main takeaways: 
PPR, run on a neighborhood graph, recovers only \emph{geometrically
compact} high-density clusters. Our main theoretical results, makes this takeaway 
precise, and provides concrete ways of quantifying the geometric
compactness of a density cluster.

@@

  We assume that the neighborhood graph radius $0 < r \leq \sigma/2d$ is chosen such that
  \begin{equation*}
  \vol_{\Pbb,r}(\mathcal{S}) \leq \frac{1}{2}\vol_{\Pbb,r}(\Reals^d).
  \end{equation*}

@@

\subsection{Cluster Recovery in Symmetric Set Difference}
We now present our first main result: a bound on the volume of the symmetric set difference between the estimated cluster $\widehat{C}$ and empirical cluster $\Csig[\Xbf]$. 
In this theorem, and hereafter, we let 
$c, c_i > 0$ represent universal constants, and 
$c_i(\Pbb,r)$ represent constants which depend only on $\Pbb$ and $r$ but not on the sample size $n$; where possible, we keep track of all constants in our proofs.
\begin{theorem}
  \label{thm:volume_ssd}
  Fix $\lambda > 0$ and let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster for some $\sigma > 0$. If Algorithm \ref{alg: ppr} is \emph{well-initialized} with respect to $\Cset$, then 
  %there exists a universal constant $c > 0$, constants in sample size $c_1(\Pbb,r)$ - $c_6(\Pbb,r)$, and 
  there exists a set $\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$ of large volume 
  with $\vol(\Csig[\Xbf]^g; G_{n,r}) \geq \frac{1}{2}\vol(\Csig[\Xbf];G_{n,r})$ such that the following statement holds: let Algorithm~\ref{alg: ppr} be run with any seed node $v \in \Csig[\Xbf]^g$. Then, for any $n$ large enough so that
  \begin{equation*}
  n \geq \max\left\{(\log n)^{\max\{\frac{3}{d},1\}}\left(\frac{1}{c_2(\Pbb,r)}\right)^d, c_3(\Pbb,r)\right\},
  \end{equation*}
  the volume of the symmetric set difference is upper bounded
  \begin{equation}
    \label{eqn:volume_ssd}
    \Delta(\Csig[\Xbf], \Cest) \leq c \cdot \kappa(\Cset) \cdot \vol_{n,r}(\Csig[\Xbf]),
  \end{equation}
  with probability at least $1 - \frac{c_4(\Pbb,r)}{n} - 2n\exp\set{-c_5(\Pbb,r)n} - 2\exp\set{-(c_1(\Pbb,r) + c_6(\Pbb,r))n}$. 
\end{theorem}
\sbcomment{At this point not sure why you need so many different constants. I would in general assume that $d$ is a constant. Can you not just write for 
$n \geq c_1(\Pbb,r)  (\log n)^{\max\{\frac{3}{d},1\}}$, with probability at least $1 - c_2(\Pbb,r)/n$ blah...?}
The proof of Theorem \ref{thm:volume_ssd} is deferred to Appendix~\ref{sec:proof_of_volume_ssd}. 
This result shows that the volume of the symmetric set difference \smash{$\Delta(\Csig[\Xbf],
  \Cest)$} is upper-bounded by a quantity proportional to the difficulty of the clustering problem, as
measured by the condition number $\kappa(\Cset)$.
\sbcomment{Not completely clear what you are doing above Zhu et. al. Perhaps emphasize again that the technical work is in showing that $\tau_u$ and $\phi_u$ are valid upper bounds on the normalized cut and mixing time, and once
this is established the result follows from Zhu et al.? (unless I am missing something) }

\subsection{Consistent Cluster Recovery}
The bound on symmetric set difference \eqref{eqn:volume_ssd} does not imply consistent density cluster estimation in the sense of \eqref{eqn: consistent_density_cluster_recovery}. This notion of consistency requires a uniform bound over 
the PPR vector $\pbf_v$: as an example, suppose that we were able to show that for all \smash{$\Cset'
  \in \Cbb_f(\lambda), \Cset' \neq \Cset$}, and each $u \in \Cset, w \in
\Cset'$,  
\begin{equation}
\label{eqn: ppr_gap}
\frac{\pbf_v(w)}{D_{ww}} \leq \frac{1}{100 {n \choose 2} \vol_{\Pbb,r}(\Csig)} < \frac{1}{10 {n \choose 2} \vol_{\Pbb,r}(\Csig)} \leq \frac{\pbf_v(u)}{D_{uu}}.
\end{equation}
Then, any $(L,U)$ satisfying~\eqref{eqn: initialization} and any sweep cut $S_{\beta}$ for $\beta \in (L,U)$ will fulfill both conditions
laid out in \eqref{eqn: consistent_density_cluster_recovery}. In Theorem
\ref{thm: consistent_recovery_of_density_clusters}, we show that a sufficiently
small upper bound on $\kappa(\Cset)$ ensures that with high probability the uniform bound~\eqref{eqn: ppr_gap} is satisfied, and hence implies \smash{$\Cest$} will be a consistent  
estimator. We will need one additional regularity condition, to preclude arbitrarily low degree
vertices for points $x \in \Cset'[\Xbf]$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{6}
\item 
  \label{asmp: C'_bounded_density}
  \emph{Bounded density in other clusters:} Letting $\sigma,\lambda_{\sigma}$ be 
  as in \ref{asmp: bounded_density}, for each $\Cset' \in \Cbb_f(\lambda)$ and
  for all $x \in \Csig'$, $\lambda_{\sigma} \leq f(x)$. 
\end{enumerate}

@@

\begin{theorem}
  \label{thm: consistent_recovery_of_density_clusters}
  Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a
  $\kappa$-well-conditioned density cluster for some $\sigma > 0$, and
  additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm
  \ref{alg: ppr} is well-initialized and if 
  %, there exists a universal constant $c >  0$ and constants in sample size $c_1(\Pbb,r)$-$c_6(\Pbb,r)$ such that if  
  \begin{equation}
    \label{eqn: kappa_ub}
    \kappa(\Cset) \leq c \frac{(\lambda_{\sigma} r^d
      \nu_d)^2}{\vol_{\Pbb,r}(\Csig)}, 
  \end{equation}
  then $\Cest$ satisfies~\eqref{eqn: consistent_density_cluster_recovery} with respect to $\Cset$ with probability at least $1 - \frac{c_4(\Pbb,r)}{n} - 2n\exp\set{-c_5(\Pbb,r)n} - 2\exp\set{-(c_1(\Pbb,r) + c_6(\Pbb,r))n}$. Therefore, $\Cest$ is a consistent estimator of $\Cset$ in the sense of Definition \ref{def: consistent_density_cluster_estimation}. 
\end{theorem}
\sbcomment{Same suggestion about removing unnecessary constants in the statement of the theorem. You can leave them in the proof.}
\noindent Some remarks are in order: 
 \begin{enumerate}  
 \item We note that the restriction on $\kappa(\Cset)$ imposed by \eqref{eqn:
    kappa_ub} results in an upper bound on the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf],
    \Cest)$} on the order of $r^d$. In plain terms, we are able to recover a
  density cluster $\Cset$ in the sense of \eqref{eqn:
    consistent_density_cluster_recovery} only when we can guarantee a very small
  fraction of points will be misclassified. This strong condition is the price
  we pay in order to obtain the uniform bound in~\eqref{eqn: ppr_gap}. 
 \item While taking the radius of the neighborhood graph $r \to 0$ as $n \to
  \infty$ (thereby ensuring that the graph $G_{n,r}$ is sparse) is computationally
  attractive, the presence of a factor of \smash{$\log^2(1/r)/r$} in
  $\kappa(\Cset)$ prevents us from making claims about the
  behavior of PPR in this regime. Although the restriction to a kernel
  function fixed in $n$ is common in spectral clustering theory
  \citep{schiebinger2015,vonluxburg2008}, it is an interesting question whether
  PPR exhibits some degeneracy over $r$-neighborhood graphs as $r \to 0$, or if
  this is merely looseness in our upper bounds.  
\end{enumerate}

\subsection{Approximate PPR vector} 
In practice, exactly solving the system of equations~\eqref{eqn:
  ppr_vector} to compute the PPR vector 
  may be too computationally expensive. To address this limitation,
\citet{andersen2006} introduced the \emph{$\varepsilon$-approximate} PPR vector
(aPPR), which we will denote by \smash{$\pbf^{(\varepsilon)}$}. We refer the
curious reader to \citet{andersen2006} for a formal algorithmic definition of
the aPPR vector, and limit ourselves to highlighting a few salient points: the
aPPR vector can be computed in order $\mathcal{O}(1/(\varepsilon\alpha))$ time,
while satisfying the following uniform error bound: 
\begin{equation}
\label{eqn: appr_error}
\textrm{for all $u \in V$}, \quad \pbf(u) - \varepsilon \Dbf_{uu}\leq
\pbf^{(\varepsilon)}(u) \leq \pbf(u).  
\end{equation}
For a sufficiently small choice of $\varepsilon$, the 
application of \eqref{eqn: appr_error} within the proofs of Theorems \ref{thm:volume_ssd} and \ref{thm: consistent_recovery_of_density_clusters}
leads to analogous results which hold for \smash{$\pbf^{(\varepsilon)}$}. We
formally state and prove this in Corollary~\ref{cor: appr} in 
Appendix~\ref{sec:appr}.
\sbcomment{Maybe we should discuss approximate PPR briefly in the Background section -- and just state the corollary here (not in a separate section).}

@@

One of the primary technical contributions of our work is showing that the geometric
conditions \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} translate to 
meaningful bounds on the normalized cut and mixing time of $\Csig[\Xbf]$ in
$G_{n,r}$. In doing so, we elaborate on how some of the geometric conditions
introduced in Section~\ref{sec: consistent_cluster_estimation_with_ppr}
contribute to the difficulty of the clustering problem.

@@

\begin{remark}Technically, the rectangles $\Csig^{(1)}$ and $\Csig^{(2)}$ are not $\sigma$-expansions due to their sharp corners. To fix this, one can 
%either modify the upper bound (specifically, Lemma~\ref{lem: expansion_volume} in the proof of Theorem~\ref{thm: conductance_upper_bound}) to hold with respect to rectangles of width $\sigma$, or 
change the sets $\Csig^{(1)}$ and $\Csig^{(2)}$ to be rectangles with rounded corners. With some additional bookkeeping, using rounded rectangles does not change our lower bound significantly so we ignore this technicality in our 
subsequent discussion.
\end{remark}

@@

\sbcomment{Sometimes you make remarks in a remark environment (so it is italicized) and sometimes not. Can you use the same format throughout? I have a preference for \emph{not} italicizing. }
